\chapter{Introduction}

Nepal is a country of extreme geographical and climatic diversity, compressed into a relatively small area. This unique combination, spanning from near sea level to the world's highest peaks, influences its climate, historical development, regional landscapes, and the lifestyles of its people. For a long, long time, Nepal's weather has decided where people settled down, what crops they grew, and even their festivals. The big summer rain, called the monsoon, is like the country's heartbeat. It brings most of the year's rain from June to September, helping farms grow food. The monsoon rains have grown more irregular and extreme over time causing floods and landslides, sometimes not enough leading to droughts. This is a big challenge, made worse by climate change affecting the whole world. The huge Himalayan mountains in the north also play a huge role. They block chilly winds from Central Asia in winter, and they catch all the monsoon rain, making some areas behind the mountains very dry

\section*{Terrain and its Influence on Climate}

Nepal's topography is arguably the most significant determinant of its climate. It's often divided into three distinct ecological belts running east to west:

\begin{enumerate}

\item \subsection*{Terai Region (Southern Lowlands)}

\textbf{Terrain:} This is a low-lying, flat, fertile plain, part of the Gangetic Plain, with elevations typically below 300 meters (around 1,000 feet). It includes some forested areas and gentle hill ranges (Siwalik Hills).\\
\textbf{Climate:} Experiences a tropical to subtropical monsoon climate. Summers are hot and humid, with temperatures often exceeding 37$^\circ$C (99$^\circ$F), and can even reach above 40$^\circ$C (104$^\circ$F) in some areas. Winters are mild, ranging from 7$^\circ$C to 23$^\circ$C (45$^\circ$F to 73$^\circ$F). The majority of rainfall occurs during the monsoon season (June to September), leading to lush vegetation but also prone to flooding.

\item \subsection*{Hilly Region (Mid-Hills)}

\textbf{Terrain:} Characterized by undulating hills, valleys (like Kathmandu and Pokhara Valleys), and ranges (Mahabharat Lekh), with altitudes generally between 800 to 4,000 meters (around 2,600 to 13,000 feet).\\
\textbf{Climate:} Features a warm temperate to cool temperate climate. Summers are generally pleasant (19$^\circ$C to 35$^\circ$C or 66$^\circ$F to 95$^\circ$F in valleys), while winters are cooler, with temperatures dropping to sub-zero at higher elevations. Kathmandu Valley has a moderate climate. This region also receives substantial monsoon rainfall, with some areas like Pokhara receiving very heavy rainfall due to the obstruction of monsoon clouds by the Annapurna range.

\item \subsection*{Himalayan Region (Northern Mountains)}

\textbf{Terrain:} Dominated by the Great Himalayan Range, including the world's highest peaks, with elevations ranging from 4,000 meters to over 8,848 meters (over 13,000 feet to 29,000 feet). This region includes high-altitude alpine valleys and vast snow-capped mountains.\\
\textbf{Climate:} Ranges from cool temperate to subarctic and arctic (or nival) climate. Summers are cool and short, while winters are severe, with temperatures consistently below freezing, often plunging to −25$^\circ$C (−13$^\circ$F) or lower at the highest altitudes. Snowfall is common, especially in winter. The Himalayas act as a massive barrier, blocking cold winds from Central Asia in winter and effectively trapping the monsoon winds from the south, leading to rain shadow areas (like Mustang and Manang) behind the main range which are arid or semi-arid.

\end{enumerate}
\section*{Lifestyle and Climate}

The diverse climate zones have directly shaped the lifestyles of the Nepali people:

\begin{enumerate}
  \item \textbf{Agriculture as a Core:} The majority of Nepal's population is engaged in subsistence agriculture, which is highly climate-dependent.
    \begin{itemize}
      \item In the Terai, tropical crops like rice, maize, and sugarcane are cultivated, benefiting from the warm, humid climate and monsoon rains.
      \item In the Hills, terraced farming is prevalent, allowing for the cultivation of a wider variety of crops suited to temperate conditions, including rice, maize, millet, wheat, and various vegetables.
      \item In the Himalayas, high-altitude communities rely on cold-tolerant crops like barley, buckwheat, and potatoes, along with pastoralism (yak herding) adapted to the harsh environment.
    \end{itemize}
  \item \textbf{Housing and Clothing:} Traditional architecture and clothing are adapted to local climatic conditions.
    \begin{itemize}
      \item Terai homes are often built with natural cooling in mind, using materials like mud, thatch, and bamboo. Lighter clothing is worn year-round.
      \item Hill homes are typically sturdier, often with stone or brick, providing insulation against cooler temperatures. Layered clothing is common.
      \item Himalayan houses are built to withstand extreme cold, using thick stone walls and often small windows. Warm, heavy woolens and animal hides are essential for survival.
    \end{itemize}
  \item \textbf{Cultural Practices and Festivals:} Many festivals and traditional practices are linked to the agricultural calendar and seasonal changes. For instance, festivals like Dashain and Tihar are celebrated after the monsoon harvest, reflecting gratitude for a successful crop.
  \item \textbf{Migration and Livelihoods:} Climate change impacts are increasingly affecting livelihoods. Unpredictable monsoons, droughts, and natural disasters are leading to reduced crop yields and food insecurity, especially in rural areas, prompting internal and external migration for alternative employment. Glacial melt threatens traditional livelihoods of Sherpa communities, forcing them to adapt or relocate.
  \item \textbf{Tourism:} The diverse climate zones also drive Nepal's tourism industry. Trekking and mountaineering are popular during the drier, clearer seasons (autumn and spring) when mountain views are optimal and temperatures are pleasant. Jungle safaris in the Terai are best during the dry winter months.
\end{enumerate}

\section*{Historical Perspective}

Historically, Nepal's climate has shaped human settlement patterns and interactions.

\begin{itemize}
  \item \textbf{Early Settlements and Trade Routes:} The fertile plains of the Terai and the milder mid-hills have long been the most welcoming places for people to live. These areas offered good soil and enough water to support farming, which helped early kingdoms and trading centers grow. The pleasant climate and reliable resources made it easier for communities to thrive and expand.

  \item \textbf{Mountain Barriers and Isolation:} The high Himalayan mountains in the north acted like a natural fortress, protecting the region but also keeping it isolated. This made it hard for outside influences to reach some mountain valleys, allowing unique cultures and traditions to develop in these remote areas. People there became largely self-reliant, shaped by the rugged environment around them.
  
  \item \textbf{Monsoon Dependence:}  For centuries, Nepal’s farming communities have depended heavily on the monsoon rains. The rhythm of planting, harvesting, and many cultural festivals has always followed the arrival of the monsoon. But when the rains are late, too light, or too heavy, it can cause serious problems like droughts, floods, or crop failures. These challenges have influenced migration patterns and the stability of communities over time.

  \item \textbf{Climate Change Impacts (Recent History):}  In recent decades, Nepal has been feeling the effects of global climate change more than most, even though it contributes very little to global emissions. Temperatures are rising, especially during the dry season, and rainfall is becoming less predictable. Winters are getting drier, while monsoon rains come in heavier bursts. These changes have led to more disasters like glacial lake floods, landslides, and severe droughts, which threaten the livelihoods of many, especially those who rely on farming and the natural environment.

\end{itemize}

\section*{Nepal’s Climatic Trends and Events}

Over the past few decades, Nepal’s climate has been changing noticeably. These changes are affecting everything from farming and water supply to wildlife and the everyday lives of people across the country.


\subsection*{Observed Climatic Trends in Nepal}

\begin{enumerate}

\item {Rising Temperatures}
\begin{itemize}
  \item According to the Department of Hydrology and Meteorology (DHM), Nepal’s average annual temperature has been increasing by approximately 0.06$^\circ$C per year since the 1970s. 
  \item The warming trend is more pronounced in the high Himalayan region compared to the lowlands. 
  \item Urban areas such as Kathmandu have experienced more frequent and prolonged heatwaves.
\end{itemize}

\item {Shifting Rainfall Patterns}
\begin{itemize}
  \item Monsoon rains, which account for over 80\% of annual precipitation, have become more erratic, with delayed onsets and sudden intense bursts. 
  \item There’s an observable increase in dry spells during the growing season and heavier rainfall over shorter periods, leading to increased risk of floods and landslides.
\end{itemize}

\item {Glacial Retreat} 
\begin{itemize}
  \item Glaciers in the Himalayas, including those feeding rivers like the Koshi and Gandaki, are retreating at alarming rates. 
  \item Studies show that more than 80\% of glaciers are shrinking, contributing to glacial lake outburst floods (GLOFs).
\end{itemize}

\item {Extreme Weather Events} 
\begin{itemize}
\item Nepal has seen a rise in extreme events like floods, landslides, droughts, and hailstorms over the past two decades. \item Changes in temperature and precipitation patterns have disrupted traditional farming calendars.
\end{itemize}
\end{enumerate}
\subsection*{Notable Climate-Related Events in Nepal}

\begin{longtable}{|p{4cm}|p{2cm}|p{8cm}|}
\hline
\textbf{Event} & \textbf{Year} & \textbf{Description} \\
\hline
\endfirsthead

\hline
\textbf{Event} & \textbf{Year} & \textbf{Description} \\
\hline
\endhead

Koshi Flood & 2008 & Breach in the Koshi embankment displaced over 70,000 people and caused massive destruction in the Terai. \\
\hline
Glacial Lake Outburst Flood (Dig Tsho) & 1985 & One of the earliest documented GLOFs in Nepal, causing loss of infrastructure and property in Khumbu. \\
\hline
Severe Drought in Mid-Western Nepal & 2009 & Affected food production and water supply, leading to a food crisis in many remote regions. \\
\hline
Melamchi Flash Flood & 2021 & Caused by intense rainfall and possible landslide dam bursts, severely damaging the Melamchi Water Project. \\
\hline
Rain-Induced Landslides & Recurring & Monsoon-triggered landslides are increasingly frequent in hill districts such as Sindhupalchok and Myagdi. \\
\hline
\end{longtable}
\captionsetup{justification=centering}
\captionof{table}{Major climate-related disasters in Nepal and their impact.}

\section*{Climate Science and Its Importance}

Climate science is the study of long-term patterns and changes in weather, temperature, and other atmospheric conditions on Earth. Unlike weather, which refers to short-term changes in the atmosphere, climate focuses on how these patterns evolve over decades, centuries, or even longer. Understanding climate science is crucial because it helps us understand the natural processes that shape our environment and how human activities, such as burning fossil fuels, affect these patterns.

Researchers in this field use sophisticated models and historical data to predict future climate changes. These predictions help in managing resources, preparing for extreme weather events, and making policy decisions to protect the planet and its inhabitants. Climate science is particularly important for several reasons:

\begin{itemize}
    \item \textbf{Human and Ecological Health:} Climate change has direct and indirect impacts on human health, including the spread of diseases, heatwaves, and the availability of clean water. Understanding climate patterns helps protect public health by predicting and mitigating these risks.
    \item \textbf{Natural Resource Management:} Climate science aids in managing natural resources like water, forests, and energy by predicting changes in the environment that might affect these resources. For instance, changes in precipitation patterns can influence water availability for agriculture and consumption.
    \item \textbf{Biodiversity Protection:} Climate science helps track how changing environmental conditions affect ecosystems and biodiversity. For example, many species are migrating or adapting to new climates due to rising temperatures, and understanding these changes is critical for protecting ecosystems.
    \item \textbf{Predicting and Preparing for Extreme Weather:} Climate scientists study the relationship between climate change and extreme weather events, such as hurricanes, floods, and wildfires. This helps in predicting these events and preparing communities, governments, and industries for their impacts.
\end{itemize}

In summary, climate science is essential for understanding the Earth’s natural systems and how human actions are influencing these systems. It provides the knowledge necessary for mitigating climate change, adapting to its effects, and making informed decisions for a sustainable future.

\section*{Data Analytics in Climate Science}

Data analytics plays a key role in climate science by transforming raw climate data into meaningful insights. Climate data is complex and often come in large volumes. Data analytics helps researchers and policy makers make sense of these data, uncover hidden trends, and make better decisions.

Here are a few ways data analytics is applied in climate science.

\begin{itemize}
    \item \textbf{Trend Analysis:} By analyzing historical climate data, scientists can identify long-term trends in temperature, precipitation, and other factors. For example, they might find that temperatures have been steadily rising in certain regions over the past 50 years.
    \item \textbf{Modeling and Predictions:} Climate models are created using statistical techniques to predict how climate will change in the future. These models use current and historical data to estimate future scenarios based on different levels of greenhouse gas emissions and other factors.
    \item \textbf{Visualization:} Data analytics tools such as graphs, maps, and charts help researchers and the public visualize complex climate data. Visualizations make it easier to understand trends, compare regions, and communicate findings.
    \item \textbf{Extreme Event Forecasting:} By analyzing patterns in weather data, data analytics can help predict extreme weather events such as hurricanes, droughts, and floods. This helps communities prepare and mitigate the effects of these events.
\end{itemize}

\input{chapter 1/eda.tex}
\input{chapter 1/packages.tex}