\section*{Do It Yourself: Scenario-Based Explorations}

In this section, you will apply your R skills to analyze real-world climate scenarios inspired by stories and local observations. Each scenario includes guiding hints and a challenge question to deepen your understanding and decision-making ability using data.

\begin{enumerate}

  \item \textbf{The Warming Winters} \\
  In recent years, winters feel shorter and less cold. Locals say, “We used to need two blankets, now one is enough.” \\
  \textbf{Hints:}
  \begin{itemize}
    \item Check the average winter temperature for each year (e.g., December to February).
    \item Plot the minimum winter temperatures over time.
    \item Compare the current winter averages to those from 10 years ago.
  \end{itemize}
  \textbf{Challenge Question:} \\
  What could happen to winter crops or water storage from snow if this warming trend continues?

  \item \textbf{The Missing Rains} \\
  A farmer says, “It rains all at once or not at all.” The rain hasn’t disappeared—but its pattern seems to have changed. \\
  \textbf{Hints:}
  \begin{itemize}
    \item Calculate how often it rains (number of rainy days) each month.
    \item Plot rainfall amounts by month over the last few years.
    \item Look for months with very heavy rainfall vs. dry periods.
  \end{itemize}
  \textbf{Challenge Question:} \\
  How can farmers plan planting if rainfall becomes less frequent but more intense?

  \item \textbf{Hotspots Emerging} \\
  Heat records are being broken in specific districts while others remain unaffected. Are climate “hotspots” forming? \\
  \textbf{Hints:}
  \begin{itemize}
    \item Calculate average summer temperatures by district.
    \item Identify districts with a steep temperature increase over the last 10 years.
    \item Rank districts by their rate of temperature rise.
  \end{itemize}
  \textbf{Challenge Question:} \\
  Label 3 emerging heat hotspots. What traits (e.g., elevation, urbanization) could explain these areas?

  \item \textbf{Wind Changes in Kathmandu} \\
  Residents have noticed changing wind patterns—some days feel breezier, others still. \\
  \textbf{Hints:}
  \begin{itemize}
    \item Calculate average wind speeds by month over multiple years.
    \item Identify any increasing or decreasing trends.
    \item Highlight the months with the strongest wind activity.
  \end{itemize}
  \textbf{Challenge Question:} \\
  How might wind changes affect air pollution or comfort? Suggest one way Kathmandu could make better use of strong winds.

  \item \textbf{Water Worries in Palpa} \\
  In Palpa, farmers are struggling with reduced and unpredictable rainfall. Water-demanding crops like paddy are suffering. \\
  \textbf{Hints:}
  \begin{itemize}
    \item Identify years in Palpa with low rainfall and high temperatures.
    \item Examine which crops had poor yields in those years.
    \item Compare the performance of rainfed and irrigated crops.
  \end{itemize}
  \textbf{Challenge Question:} \\
  What alternative crops or farming methods would you recommend for Palpa’s farmers during dry conditions? Support your answer using data.

\end{enumerate}
