\chapter{Merging with Agriculture Dataset}

Imagine you are a farmer or an agricultural planner trying to make the best decisions for the upcoming planting season. What if you could peek into the climate’s behavior — temperature changes, rainfall patterns, wind speed — and understand exactly how these factors affect crop growth?

In this chapter, we take a fascinating step forward: we merge agricultural data with climate data. By combining these two worlds, we can uncover deeper insights about how weather and climate influence agricultural productivity, helping farmers, scientists, and policymakers make smarter, data-driven decisions. Let’s think about this together:

\begin{itemize}
    \item How do you think temperature fluctuations affect crop yield?
    \item What role does rainfall play in the growth cycle of different crops?
    \item Can wind patterns influence soil erosion or pollination?
\end{itemize}

By the end of this chapter, you’ll not only learn how to merge these datasets technically, but also explore why this integration is so crucial for understanding the bigger picture of agriculture under changing climate conditions.

Ready to see the magic happen when data meets reality? Let’s dive in!

\subsection*{Sourcing Agricultural Data}
To carry out this integration, we sourced agricultural data from a publicly available interactive visualization platform: \\
\href{https://public.tableau.com/app/profile/sadichchha1369/viz/NepaCropMapwithprovinceSAMPLE_15636442141840/Dashboard1}{Tableau Public - Nepal Crop Map Dashboard}\\
This dataset includes district-level agricultural information, which we align with our climate dataset to explore temporal and spatial patterns in greater depth.
But before merging, let’s take a step back. Data, especially from real-world sources is rarely perfect. It often contains missing entries, inconsistent formats, or irrelevant information. That’s why our first task is to clean the agricultural dataset to ensure it aligns smoothly with our climate data.


\input{chapter 5/prep}
\input{chapter 5/merge}
\input{chapter 5/visualise}

