\section*{TABLE C: R Packages with their Functionalities}

\begin{longtable}{|p{4cm}|p{3cm}|p{7cm}|}
\hline
\textbf{Functionality} & \textbf{Package Name} & \textbf{Description / Use} \\
\endfirsthead
\hline
\textbf{Functionality} & \textbf{Package Name} & \textbf{Description / Use} \\
\endhead
\hline
\textbf{Data Wrangling} & dplyr & Grammar of data manipulation: filter, mutate, select, arrange, summarize. \\
\hline
 & tidyr & Data tidying: pivot\_longer, pivot\_wider, separate, unite. \\
\hline
 & data.table & Fast data manipulation, especially for large datasets. \\
\hline
 & readr & Fast reading of CSV, TSV, and other flat files. \\
\hline
 & stringr & Consistent and easy string manipulation functions. \\
\hline
 & lubridate & Makes working with dates and times easy and readable. \\
\hline
\textbf{Data Visualization} & ggplot2 & Widely used for elegant and layered data visualizations. \\
\hline
 & plotly & Interactive plots built on top of ggplot2 or independently. \\
\hline
 & lattice & Powerful system for creating trellis graphs. \\
\hline
\textbf{Statistical Modeling} & stats (base R) & Built-in statistical tests, linear models, hypothesis testing. \\
\hline
 & caret & Unified interface for training, tuning, and evaluating ML models. \\
\hline
 & glmnet & Regularized regression: lasso and ridge regression. \\
\hline
 & randomForest & Implements random forest classification and regression. \\
\hline
 & xgboost & Extreme gradient boosting: powerful ML model for structured data. \\
\hline
 & nnet & Feed-forward neural networks and multinomial log-linear models. \\
\hline
\textbf{Time Series Analysis} & forecast & Functions for forecasting and time series modeling (ARIMA, ETS, etc.). \\
\hline
 & tseries & Time series analysis including unit root tests and GARCH models. \\
\hline
 & tsibble & Tidy temporal data frames (for use with fable). \\
\hline
 & zoo / xts & Time series objects and manipulation. \\
\hline
\textbf{Reporting} & rmarkdown & Dynamic report generation in HTML, PDF, Word, etc. \\
\hline
 & knitr & Integrates R code chunks into LaTeX, Markdown, or HTML. \\
\hline
\textbf{Spatial Analysis / Mapping} & sf & Simple features for working with geospatial data. \\
\hline
 & tmap & Thematic maps based on the Grammar of Graphics. \\
\hline
 & leaflet & Interactive maps using Leaflet.js. \\
\hline
 & sp / rgdal / rgeos & Classic spatial data handling and GIS integration. \\
\hline
\textbf{Machine Learning} & mlr3 & Object-oriented framework for ML tasks, workflows, and benchmarking. \\
\hline
 & tidymodels & Tidy approach to ML using parsnip, recipes, and tune. \\
\hline
 & h2o & Scalable ML with deep learning, GBM, random forests, etc. \\
\hline
\textbf{Shiny Apps \& Dashboards} & shiny & Create interactive web apps directly from R. \\
\hline
 & shinydashboard & Structured layout system for building dashboards in Shiny. \\
\hline
 & shinyWidgets & Enhancements to basic Shiny UI components. \\
\hline
 & flexdashboard & Create interactive dashboards with R Markdown. \\
\hline
\end{longtable}
% \caption{R Packages and Their Functionalities}
\clearpage

